---
title: "hw7"
author: "Kamala Dadashova"
date: "3/4/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

 $\textbf{Problem 4:}$ The mileage data sample from Example 7.6 in Devore is contained in a comma separated
value file called mileage.csv. Use this data to do the following in R:


 
(a)  Make a Q-Q plot of the data in the sample. Based on your observations, can we assume this data came from a normal distribution?

```{r qqplot}

mileages <- read.csv("~/Desktop/Statistics Learn/ST515-2022/WEEK7/Confidence Intervals-20220301/HW7/mileage.csv")
mile = mileages$mileage
n = length(mile)
qqnorm(mile, pch=19)
qqline(mile, lwd=2, col="blue")
```

As we observed, data has a normal distribution. 

(b) Calculate a large sample approximate confidence interval for the mean and an exact confidence interval for the mean based on the assumption the data is normally distributed. Compare the results. What does this say about your conclusion in (a). ?
```{r confidence}
# Large sample approximation.
approx <- c(mean(mile)-1.96*sd(mile)/sqrt(n),mean(mile)+1.96*sd(mile)/sqrt(n))
names(approx) = c("lower","upper")
approx

#Exact.
t.crit <- qt(0.025,n-1,lower.tail=F)
exact <- c(mean(mile)-t.crit*sd(mile)/sqrt(n),mean(mile)+t.crit*sd(mile)/sqrt(n))  
names(exact) = c("lower","upper")
exact
```
In terms of comparison, we observe that the there is a small difference in lower bounds and in upper bounds for both large and exact sample approximation implying that our assumption in part (a) about normal distribution holds true. 





(c) Generate a bootstrap percentile confidence interval for the mean from the sample
data.

```{r bootstrap}

B <- 10000  # run 10000 bootstrap samples
means <- rep(0,B)  # create dummy matric for means
for(i in 1:B){
  rsample <- sample(mile,n,replace=T) 
  means[i] <- mean(rsample)  # calculate mean and save in vector
}
x_bar_bar <- mean(means)  # calculate mean estimate


# t-CI with bootstrap SE
tboot <- c(mean(mile)-qt(0.025,n-1,lower.tail=F)*sd(means),
           mean(mile)+qt(0.025,n-1,lower.tail=F)*sd(means))
names(tboot) = c("lower", "upper")
tboot

# bootstrap percentile confidence interval
pctile <- quantile(means,probs=c(0.025,0.975))
pctile
```




(d) Which interval is most appropriate for this data? Justify your answer.

```{r}
# plot and compare
par(mfrow=c(1,1))
hist(means,breaks=20,col="light blue")  #histogram of means
abline(v=exact,lty=1)   # overlay exact interval
abline(v=tboot,lty=2)   # overlay t-bootstrap interval
abline(v=pctile,lty=3)  # overlay percentile interval
abline(v=approx,lty=5)  # overlay percentile interval
legend("topleft",legend=c("exact","tboot","pctile",'approx.'),lty=c(1,2,3,5),bty="n")

```

 As we observe from histogram, tboot gives the best approximation to exact value.
 

 $\textbf{Problem 5:}$ Data from a sample of maximum pressures on concrete framework is contained in the file
pressure.csv. Use this data to do the following in R:



(a) Make Q-Q plot of the data in the sample. Based on your observations, can we assume this data came from a normal distribution?
```{r}
pressure <- read.csv("~/Desktop/Statistics Learn/ST515-2022/WEEK7/Confidence Intervals-20220301/HW7/pressure.csv")
pressure = pressure[,1]
n = length(pressure)
qqnorm(pressure, pch=19)
qqline(pressure, lwd=2, col="blue")
```
 
 As we observe, the data is normally distributed.


(b) Calculate an exact confidence interval for both the mean and the standard deviation based on the assumption the data is normally distributed.

```{r}
#Exact.
t.crit <- qt(0.025,n-1,lower.tail=F)
exact <- c(mean(pressure)-t.crit*sd(pressure)/sqrt(n),mean(pressure)+t.crit*sd(pressure)/sqrt(n))  
names(exact) = c("lower","upper")
exact

X2l.crit <- qchisq(0.025,n-1,lower.tail=T)
X2u.crit <- qchisq(0.975,n-1,lower.tail=T)
sd_interval <- c(sqrt((n-1)*var(pressure)/X2u.crit),sqrt((n-1)*var(pressure)/X2l.crit))
names(sd_interval)<-c("lower", "upper")
sd_interval
```


(c) Generate a bootstrap percentile confidence interval for the mean and the standard deviation from the sample data.

```{r}

B <- 10000  # run 10000 bootstrap samples
means <- rep(0,B)  # create dummy matric for means
for(i in 1:B){
  rsample <- sample(pressure,n,replace=T) 
  means[i] <- mean(rsample)  # calculate mean and save in vector
}
x_bar_bar <- mean(means)  # calculate mean estimate
se_est <- sd(means) #standard error estimates
se_est
# t-CI with bootstrap SE
tboot <- c(mean(pressure)-qt(0.025,n-1,lower.tail=F)*sd(means),
           mean(pressure)+qt(0.025,n-1,lower.tail=F)*sd(means))
names(tboot) = c("lower", "upper")
tboot

# bootstrap percentile confidence interval
pctile <- quantile(means,probs=c(0.025,0.975))
pctile
```




(d) Which intervals are most appropriate for this data? Justify your answer
 
 
```{r}
# plot and compare
par(mfrow=c(1,1))
hist(means,breaks=20,col="light blue")  #histogram of means
abline(v=exact,lty=1)   # overlay exact interval
abline(v=tboot,lty=2)   # overlay t-bootstrap interval
abline(v=pctile,lty=3)  # overlay percentile interval
legend("topleft",legend=c("exact","tboot","pctile"),lty=c(1,2,3),bty="n")

```
 
 As we observe from histogram, tboot gives the best approximation to exact value.
 
 
 